---
title: "Assignment_2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(arules)
library(arulesViz)
```

## Practice with association rule mining

Question: Pick your own thresholds for lift and confidence; just be clear what these thresholds are and how you picked them. Do your discovered item sets make sense? Present your discoveries in an interesting and concise way.

Answer: 
```{r}
grocery <- read.transactions('groceries.txt', sep=',')
summary(grocery)
inspect(grocery[1:5])

itemFrequencyPlot(grocery, topN = 20) #look at frequent items 

rules <- apriori(grocery, parameter=list(support=0.01, confidence=0.5, maxlen=6))
inspect(rules[1:5])

inspect(subset(rules, subset=confidence > 0.8))
summary(rules)


```

Support=Number of transactions with both A and B / Total number of transactions=P(A∩B)

Confidence=Number of transactions with both A and B / Total number of transactions with A=P(A∩B) / P(A)

ExpectedConfidence=Number of transactions with B / Total number of transactions=P(B)

Lift=Confidence / Expected Confidence=P(A∩B) / P(A)*P(B)

## Visualizations

```{r}
rules_sorted <- sort(groceryrules, by='confidence', decreasing=TRUE)
#matrix representation
plot(groceryrules[1:20], method = 'matrix', control = list(reorder=TRUE))
#Interactive Scatterplot 
plotly_arules(groceryrules)

plot(groceryrules, method = 'graph', interactive=TRUE, shading=NA)
subrules <- head(sort(groceryrules, by='lift'),10) #Graph just 10 rules by 10 highest lifts 
plot(subrules, method='graph')
plot(groceryrules, method='grouped') #Grouped Matrix shows LHS and RHS 
plot(subrules,method='paracoord', control=list(reorder=TRUE)) #Parallel Coordinates plot for 10 rules 

```
People tend to buy semi.finished.breads and margarine before ready soups 
so we should put those items close to the ready soups isle. 
Also, people tend to buy citrus fruit, soda, canned/bottled beer, and shopping bags before they buy margarine and ready.soups. 
When they buy whole milk together with ready soups, they tend to buy other vegetables.

# How to get the product recommendation rules?

```{r}
rules_conf <- sort(rules, by='confidence', decreasing=TRUE)
inspect(head(rules_conf)) #High-confidence rules

rules_lift <- sort(rules, by='lift', decreasing=TRUE)
inspect(head(rules_lift)) #High lift rules 
```
The rules with confidence of 1 imply that, whenever the LHS item was purchased, the RHS item was also purchased 100% of the time.
So in our grocery rules, if one buys citrus fruit and root vegetables, there's 58.6% chance they will buy other vegetables. 

A rule with a lift of 3 imply that, the items in LHS and RHS are 3 times more likely to be purchased together compared to the purchases when they are assumed to be unrelated, which is for the same LHS-RHS pair of {citrus fruit, root vegetables} -> {other vegetables}.

# Targeting Items 
- what are customers likely to buy before or after this item
What are people buying before they buy ready soups?

Tend to buy citrus.fruit, kitchen towels. Consider when planning isles 

```{r}
rules <- apriori(data=grocery, parameter=list(supp=0.001, conf=0.08), appearance = list(default = 'lhs', rhs = 'ready.soups='), control=list(verbose=F))
rules <- sort(rules, decreasing=TRUE, by='confidence')
inspect(rules[1:5])
```
What are people buying after they buy ready soups? 
```{r}
rules2 <- apriori(data=grocery, parameter=list(supp=0.001, conf=0.08), appearance = list(default = 'rhs', lhs = 'ready.soups='), control=list(verbose=F))
rules2 <- sort(rules2, by='confidence', decreasing=TRUE)
inspect(rules2)
```
They tend to buy margarine and semi.finished.bread before ready soups. 

##Interactive inspect with datatable
```{r}
rules3 <- apriori(grocery, parameter=list(support=0.005, confidence=0.5))
rules4 <- apriori(grocery, parameter=list(support=0.05, confidence=0.5))
inspectDT(rules3) #if we want highest support 9 pairs 
inspectDT(rules4) #if we want wider association with 170 entries 
```


